---
title: "Figures for DigiThrace Conference proceedings"
author: "Adela Sobotkova"
date: "2024-06-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE)
library(sf)
library(raster)
library(tidyverse)
library(mapview)
```

This script ties allows for the regeneration of all figures used in the Intervisibility paper for the DigitalThrace Conference Proceedings from April 2024 meeting in Sofia.


```{r load-data}
# Terrain BOM
Y_elev <- raster("data/large/Yelev32635.tif")
plot(Y_elev)
# Yambol mounds
Yam_mnds <- readRDS("data/Yam_dd_mnds.rds") # 1073 features
```

```{r prep-maps}
# Yambol data
Y_vis <- readRDS("output_data/Yam_mnds_intervis.rds")
BG <- read_sf("data/Bulgaria_Border.shp")
region <- read_sf("data/YamRegion.shp")

# Yambol rasters
library(terra)
Y <- rast("data/large/Y_elev_ext.tif")
d <- st_make_grid(st_bbox(st_buffer(region, 5000)), n =1)
dem_Y5 <- crop(Y, d)
```

## Figure 01 - Mound chronology

Run script 05_Chronology.rmd in order to download and join the Izvestiya and AOR data.

Temporal span of the 295 burial assemblages for the 40 excavated mounds in Yambol 1960 -2014.
```{r fig1, eval = FALSE}
tiff( "figures/01.tiff", width = 8, height = 5, units = "in", res = 300)
all %>% 
  filter(!is.na(StartDate)) %>% 
  pull(StartDate) %>% table() %>% plot(ylab = "Grave count", main = "Temporal span of 295 burial assemblages from 40 mounds in Yambol")
dev.off()
```


## Figure 02 - Raster profile
```{r raster-profile-fig2}
source("scripts/VisibilityFunctions.R")

profile <- function(IDorigin, IDtarget){
     # must be an elevation raster
  # library(raster)
  # if(!exists(Y_elev)){
  #    Y_elev <- raster("../output_data/large/Yelev32635.tif")}
     # must be a simple feature with same CRS as raster
  library(sf) 
  # target just as a dataframe of x, y, WITHOUT height
  origin <- Yam_mnds %>%  
         filter(TRAP == IDorigin) %>% 
          st_coordinates()  
  target <- Yam_mnds %>% 
         filter(TRAP == IDtarget) %>%
          st_coordinates()
  # target needs to be a dataframe of x, y, and height
  testprofile <- rasterprofile(Y_elev,  origin, target)
  x = c(0,nrow(testprofile))
  y = testprofile$z[c(1,nrow(testprofile))]

    # Plot profile with start and end point elevation
  plot(testprofile$z, type = "l", 
       main = paste0("Terrain profile between mounds ", IDorigin, " and ", IDtarget),  
       ylab = "Elevation (m)", xlab = "Distance in raster pixels (1 pixel = 30m)"); 
  lines(x,y, col = "red", lwd = 2)
  
}

 # these two points are at opposite sides of the region

profile(9412, 8242)
profile(9044, 8251)

# Print figure
tiff("figures/02.tiff", width = 9, height = 6, units = "in", res = 300)
profile(9412, 8242)
dev.off()
```


## Figure 03 - Intervisibility Histograms for BOM, Veg10, Veg20
 
```{r hist-fig3}

intervis <- readRDS("output_data/Y_intervis_class_vegsim.rds")

# faceted
class_names <- c(`max_class` = "Bare-earth model",
                    `v10_class` = "Static 10m vegetation",
                    `v20_class` = "Variable 20m vegetation"
                    )

intervis %>%
  pivot_longer(cols = max_class:v20_class, 
               names_to = "class_type",
                 values_to = "class") %>% 
  ggplot() +
  geom_histogram(aes(x = class,  fill = class_type), stat="count", show.legend = FALSE ) +
  facet_wrap(~class_type,
             labeller = as_labeller(class_names))+
  theme_bw()+
  scale_fill_grey(start = .2, end = .8)+
  labs(x = "The impact of vegetation: How many other mounds can a mound see?")  # add text to the columns?

# ggsave("figures/vegmodel_histogram_bw.png", width = 7, height = 4)
ggsave("figures/03.tiff", width = 7, height = 4, dpi = 300, units = "in")
```

## Figure 04 - Winner 9412

```{r prep-fig4}
visiblefrom9412 <- Y_vis %>% 
  filter(TRAPorigin == 9412, visibility == 1) %>% 
  arrange(desc(e)) %>% 
  distinct(TRAPtarget) %>% 
  pull() 

mostvisiblemnd <- visiblefrom9412 # fill in model (veg10, veg20, none) for which we are calculating the lines of sight to the most visible mound

coords <- as.matrix(cbind(as.data.frame(st_coordinates(Yam_mnds %>% filter(TRAP == 9412))),
                as.data.frame(st_coordinates(Yam_mnds %>% filter(TRAP %in% mostvisiblemnd)))))
coords_out <- as.matrix(cbind(as.data.frame(st_coordinates(Yam_mnds %>% filter(TRAP == 9412))),
                as.data.frame(st_coordinates(mm_Yext %>% filter( Id %in% mostvisiblemnd)))))
coords <- rbind(coords, coords_out)

lines_sm <-  st_sfc(
     lapply(1:nrow(coords),
           function(i){
             st_linestring(matrix(coords[i,],ncol=2,byrow=TRUE))
           }))

st_crs(lines_sm) <- st_crs(Yam_mnds)


library(mapview)
mapview(lines_sm)+ 
  mapview(Yam_mnds %>% filter(TRAP %in% mostvisiblemnd))
```


```{r print-fig4}
# Start with map elements designed for 10in map offprint
# Function to add a north arrow
add_north_arrow <- function(x, y, length = 1, label = "N") {
  arrows(x, y, x, y + length, col = "black", lwd = 2)
  text(x, y + length + 1000, label, cex = 1, col = "black")
}

# Function to add a scale bar
add_scale_bar <- function(x, y, length = 25000, label = "25 km") {
  segments(x, y, x + length, y, col = "black", lwd = 2)
  segments(x, y, x, y + 1000, col = "black", lwd = 2)
  segments(x + length, y, x + length, y + 0.2, col = "black", lwd = 2)
  text(x + length / 2, y - 1000, label, cex = 1, col = "black")
}

tiff("figures/05A.tiff", width = 10, height = 10, units = "in", res = 300)
plot(dem_Y5, col = grey.colors(5, start = 1, end = 0)); plot(region$geometry, add =T); plot(Yam_mnds$geometry, cex= 0.4, add =T); plot(mm_Yext$geometry, cex = 0.4, add =T);
# plot(Y_rivers, col= "blue", add =T)
plot(BG$geometry, add = T, lwd = 4); plot(lines_sm, add =T)
# Add north arrow at coordinates (8, 8)
add_north_arrow(502000, 4642000, length = 2500, label = "N")
# Add scale bar at coordinates (2,502000, 4645000 1)
add_scale_bar(480000, 4640000)

# to add letter A to top right corner
text(x = par("usr")[2] - 0.1 * (par("usr")[2] - par("usr")[1]), 
     y = par("usr")[4] - 0.05 * (par("usr")[4] - par("usr")[3]), 
     labels = "A", 
     cex = 3, # Size of the label
     font = 1) # Bold text

dev.off()
```

## Figure 05 - Winner 8007
```{r prep-maps}
# Yambol data
Y_vis <- readRDS("output_data/Yam_mnds_intervis.rds")
BG <- read_sf("data/Bulgaria_Border.shp")
region <- read_sf("data/YamRegion.shp")
mm_Yext <- read_sf("output_data/Y25_mapmounds.shp")

# Yambol rasters
library(terra)
Y <- rast("data/large/Y_elev_ext.tif")
d <- st_make_grid(st_bbox(st_buffer(region, 5000)), n =1)
dem_Y5 <- crop(Y, d)
```

```{r prep-fig05}

Y8007 <- readRDS("output_data/Y8007_intervis.rds")
visiblefrom8007 <- Y8007 %>% 
  filter(visibility == 1) %>% 
  pull(TRAPtarget)
mostvisiblemnd <- visiblefrom8007 # fill in model (veg10, veg20, none) for which we are calculating the lines of sight to the most visible mound

coords <- as.matrix(cbind(as.data.frame(st_coordinates(Yam_mnds %>% filter(TRAP == 8007))),
                as.data.frame(st_coordinates(Yam_mnds %>% filter(TRAP %in% mostvisiblemnd)))))
coords_out <- as.matrix(cbind(as.data.frame(st_coordinates(Yam_mnds %>% filter(TRAP == 8007))),
                as.data.frame(st_coordinates(mm_Yext %>% filter( Id %in% mostvisiblemnd)))))
coords <- rbind(coords, coords_out)


lines_8007 <-  st_sfc(
     lapply(1:nrow(coords),
           function(i){
             st_linestring(matrix(coords[i,],ncol=2,byrow=TRUE))
           }))

st_crs(lines_8007) <- st_crs(Yam_mnds)


library(mapview)
mapview(lines_8007)+ 
  mapview(Yam_mnds %>% filter(TRAP %in% mostvisiblemnd))

```

```{r}
tiff("figures/05.tiff", width = 10, height = 10, units = "in", res = 300)
plot(dem_Y5, col = grey.colors(5, start = 1, end = 0)); plot(region$geometry, add =T); plot(Yam_mnds$geometry, cex= 0.4, add =T); plot(mm_Yext$geometry, cex = 0.4, add =T);
# plot(Y_rivers, col= "blue", add =T)
plot(BG$geometry, add = T, lwd = 4); plot(lines_8007, add =T)
# Add north arrow at coordinates (8, 8)
add_north_arrow(502000, 4642000, length = 2500, label = "N")
# Add scale bar at coordinates (2,502000, 4645000 1)
add_scale_bar(480000, 4640000)

dev.off()
```

```{r}
tiff("figures/05B.tiff", width = 10, height = 10, units = "in", res = 300)
plot(dem_Y5, col = grey.colors(5, start = 1, end = 0)); plot(region$geometry, add =T); plot(Yam_mnds$geometry, cex= 0.4, add =T); plot(mm_Yext$geometry, cex = 0.4, add =T);
# plot(Y_rivers, col= "blue", add =T)
plot(BG$geometry, add = T, lwd = 4); plot(lines_8007, add =T)
# Add north arrow at coordinates (8, 8)
add_north_arrow(502000, 4642000, length = 2500, label = "N")
# Add scale bar at coordinates (2,502000, 4645000 1)
add_scale_bar(480000, 4640000)

text(x = par("usr")[2] - 0.1 * (par("usr")[2] - par("usr")[1]), 
     y = par("usr")[4] - 0.05 * (par("usr")[4] - par("usr")[3]), 
     labels = "B", 
     cex = 3, # Size of the label
     font = 1) # Bold text
dev.off()

```


