---
title: "Mound chronology"
author: "Adela Sobotkova"
date: "2024-06-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE)
library(sf)
library(raster)
library(tidyverse)
library(mapview)
```

## Mound data
```{r load-data}
# Terrain BOM
Y_elev <- raster("data/large/Yelev32635.tif")

# Yambol mounds
Yam_mnds <- readRDS("data/Yam_dd_mnds.rds") # 1073 features
```

## Mound Chronology
Downloading the mound and burial data for each of the two literature corpus datasets, Izvestia and AOR. Now streamlined with Toshko's and Angel's help.
```{r aor-izv-google}
library(googlesheets4)
gs4_deauth()
mnd_izv <- read_sheet("https://docs.google.com/spreadsheets/d/1wOxbKVHGNHox4h86Z5ZXXADUubMJofC1KqDLMLT13a8/edit#gid=369527838", 
                      range = "General",
                      col_types = 'nccccnnccncccccccccccc')

burial_izv <- read_sheet("https://docs.google.com/spreadsheets/d/1wOxbKVHGNHox4h86Z5ZXXADUubMJofC1KqDLMLT13a8/edit#gid=369527838", 
                      range = "BurialAttributes",
                      col_types = 'ncnccccnccccccccccccccccccccnncccc')

mnd_aor <- read_sheet("https://docs.google.com/spreadsheets/d/1cx0nntcCLgrwQvCvvIYkjFJ-TnoJ0QRJfTgrf2qEEv4/edit#gid=1795672216", 
                      range = "GeneralSpatial",
                      col_types = 'nccnnncccc')

burial_aor <- read_sheet("https://docs.google.com/spreadsheets/d/1cx0nntcCLgrwQvCvvIYkjFJ-TnoJ0QRJfTgrf2qEEv4/edit#gid=1795672216", 
                      range = "BurialAttributes",
                      col_types = 'nccccnccccccccccccccnnc')

# glimpse(mnd_izv)
# glimpse(burial_izv)
# glimpse(mnd_aor)
# glimpse(burial_izv)

Yam_aor <- mnd_aor %>% 
  full_join(burial_aor, by = "MoundID") %>% 
  filter(!is.na(Long)) %>% 
  st_as_sf(coords = c("Long", "Lat"), crs = 4326) %>% 
  filter(Region == "Yambol") 
 
unique(Yam_aor$MoundID) # 40 mounds

Yam_izv <- mnd_izv %>% 
  full_join(burial_izv, by = "MoundID") %>% 
  filter(!is.na(Long)) %>% 
  st_as_sf(coords = c("Long", "Lat"), crs = 4326) %>% 
  filter(Region == "Yambol")

unique(Yam_izv$MoundID) # 10 mounds

```

## Plot the AOR and IZV mounds's start date
I select only mounds with dated cultural material, and plot their coordinates.
The, plotting all the grave's assemblages' start-dates I generate a chronology overview.
```{r plot-chrono}
Yam_aor %>% 
   filter(!is.na(StartDate)) %>% 
  #distinct(MoundID) %>% 
  mapview()
  
Yam_izv %>% 
  filter(!is.na(StartDate)) %>% 
  #distinct(MoundID) %>% 
  mapview()
Yam_izv$StartDate
Yam_izv$Enddate


# Which attributes overlap
names <- names(Yam_aor)[which(names(Yam_aor)%in%names(Yam_izv))]

all <- rbind(Yam_aor[names], Yam_izv[names])

all %>% 
  filter(!is.na(StartDate)) %>% # 295 graves
  # distinct(MoundID)  #41 mounds
  mapview(zcol = "StartDate")

# Chronology span for burial mound use
all %>% 
  filter(!is.na(StartDate)) %>% 
  pull(StartDate) %>% table() %>% plot(ylab = "Grave count", main = "Temporal span of 295 burial assemblages from 40 mounds in Yambol")

```

## Figure 01 - for printing
Temporal span of the 295 burial assemblages for the 40 excavated mounds in Yambol 1960 -2014.
```{r eval = FALSE}
tiff( "figures/01.tiff", width = 8, height = 5, units = "in", res = 300)
all %>% 
  filter(!is.na(StartDate)) %>% 
  pull(StartDate) %>% table() %>% plot(ylab = "Grave count", main = "Temporal span of 295 burial assemblages from 40 mounds in Yambol")
dev.off()
```

## Where are the mounds with dates located?

```{r map-chrono}
mapview(Yam_aor,zcol = "StartDate" ) + mapview(Yam_izv, zcol = "StartDate") + 
  mapview(Yam_mnds, cex = 1)
mapview(Yam_mnds %>% filter(prom250mbuff >75))
mapview(Yam_aor %>% filter(MoundID == 1476)) # additional review on 16 July 2024 fixed Western necropolis and Golyam Dervent dolmen which was wrongly digitized.

mapview(all,zcol = "StartDate" ) +   mapview(Yam_mnds, cex = 1)
mapview(all,zcol = "Enddate" ) +   mapview(Yam_mnds, cex = 1)

```
