---
title: "Visibility"
author: "Adela Sobotkova"
date: "2022-12-14"
output: html_document
---
## Purpose and prerequisites
This script calculates mutual intervisibility (line of sight) between various archaeological features, starting here with burial mounds in the Yambol Region within ASTER elevation raster (30m spatial resolution). 

This markdown guides you to: 

1. Define functions to calculate line-of-sight visibility between two features and automate the process for a 1000 features. 
2. Visualize intervisibility for calculated features (here: mounds in Yambol). To visualize, you can skip directly to section "Visualizing Intervisibility".
3. TBD: intervisibility between mounds and settlements

The script basically works, calculating linestrings, raster profiles, and intervisibility between Yam_mounds and generating maps

The proof of concept is done on BA mounds and to replicate it you might need: 
- to run 09 BA mounds to have the necessary libraries and digital objects

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE)
library(sf)
library(raster)
library(tidyverse)
library(mapview)
```

## Load 2009-2022 data
```{r load-mounds}

Y_elev <- raster("data/large/Yelev32635.tif")
# Yambol mounds
Yam_mnds <- readRDS("data/Yam_dd_mnds.rds") # 1073 features
mapview(Yam_mnds, zcol = "Type")

Yam_mnds %>% 
  group_by(Type) %>% 
  tally()
```
### Prominence
```{r}
mnd_85 <- Yam_mnds %>% 
  dplyr::filter(prom250mbuff>85,
                HeightMax > 1.5) 
  
mapview(Yam_mnds, cex = 0.1)+
mapview(mnd_85, cex = "prom250mbuff", zcol = "HeightMax")
```

 
## Intervisibility functions

Intervisibility in human terms mean that two or more points in the landscape have a clear line of sight between them (given reasonable visibility). In computational terms, it means that a straight line connecting the z values of the A and B (the two end-points) does not intersect a polygon formed by terrain profile between them. If it doesn't, A can see B.
Let's test such line of sight by extracting data out of raster cells between our BA points.

https://stackoverflow.com/questions/21841387/r-code-that-evaluates-line-of-sight-los-between-two-lat-lon-points 

https://gis.stackexchange.com/questions/272122/performing-viewshed-analysis-in-r
```{r define-cansee-functions}
cansee <- function(r, xy1, xy2, h1=0, h2=0){
### can xy1 see xy2 on DEM r?
### Y_elev is a DEM in same x,y, z units
### xy1 and xy2 are 2-length vectors of x,y coords
### h1 and h2 are extra height offsets (ie. mound heights)
###  (eg top of mast, observer on a ladder etc)
    xyz = rasterprofile(r, xy1, xy2)
    np = nrow(xyz)-1
    h1 = xyz$z[1] + h1
    h2 = xyz$z[np] + h2
    hpath = h1 + (0:np)*(h2-h1)/np
    return(!any(hpath < xyz$z))
}

viewTo <- function(r, xy, xy2, h1=0, h2=0, progress="none"){
    ## xy2 is a matrix of x,y coords (not a data frame)
    require(dplyr)
    apply(xy2, 1, function(d){cansee(r,xy,d,h1,h2)}, .progress=progress)
}

viewTo <- function(r, xy, xy2, h1=0, h2=0){
    ## xy2 is a matrix of x,y coords (not a data frame)
    require(dplyr)
    apply(xy2, 1, function(d){cansee(r,xy,d,h1,h2)})
}

rasterprofile <- function(r, xy1, xy2){
### sample a raster along a straight line between two points
### try to match the sampling size to the raster resolution
    dx = sqrt( (xy1[1]-xy2[1])^2 + (xy1[2]-xy2[2])^2 )
    nsteps = 1 + round(dx/ min(res(r)))
    xc = xy1[1] + (0:nsteps) * (xy2[1]-xy1[1])/nsteps
    yc = xy1[2] + (0:nsteps) * (xy2[2]-xy1[2])/nsteps
    data.frame(x=xc, y=yc, z=r[cellFromXY(r,cbind(xc,yc))])
}
```

Let's test these three functions with real data, first the `cansee()`


## Automating the intervisibility calculation

### prep the data for cansee() function
```{r auto-prep}
# origin point coordinates and height
Yam_mnds %>% 
  group_by(Type) %>% tally()

Yam_mnds$HeightMax[is.na(Yam_mnds$HeightMax)] <- 0


origin <- cbind(st_coordinates(Yam_mnds), h = Yam_mnds$HeightMax)
target <- origin 

class(origin)
```

### slow automation via a for-loop
Loops take long so evaluation is off.
```{r loop, eval = FALSE}
# LOOP: works but takes days to finish, reduce the i and e if running a test!
result_table <- NULL
for (i in 1:nrow(origin)){
  for (e in 1:nrow(target)){print(e)
      result = cansee(Y_elev, origin[i,1:2], target[e,1:2], h1 = origin[i,3], h2 = target[e,3])
    result_row <- cbind(i,e, result) 
    result_table <- rbind(result_table, result_row)
  }

}

# I stopped the process after 24 hours when it reached ~250,000 calculations
# maxima were i = 241, and e = 780 ; restart the calculations there

head(result_table)
tail(result_table)
saveRDS(result_table, "../output_data/Yam_mnd_intervis250k.rds")

```

Finishing the 1,000,000 calculations takes a while! (1 hr so far at 17.18 on Thursday night).
OK, some 24 hours later (Friday afternoon), 241 mounds (*1000) were completed. Clearly the tasks takes 4 days unless we successfully ...

### Parallelize!
Parallelisation is much faster than loops but still takes ca 6-12 hours for the whole dataset so run it on demand, if needed
```{r auto-foreach, eval = FALSE}
# install.packages("doParallel")
# install.packages("foreach")
library(foreach)
library(doParallel)
library(data.table)

detectCores() # 22 as some are multi-threaded
detectCores(logical = FALSE) # 16 real ones

cl <- makeCluster(12) # keep 4 cores so screensaver can run on W11
registerDoParallel(cl)

# Start experiment
origin <- cbind(st_coordinates(Yam_mnds), h = Yam_mnds$HeightMax)
target <- origin 

# nesting both i and e with foreach 
ie_table <- foreach (i= 1:nrow(origin), .combine = 'rbind',.packages = c("data.table", "raster")) %do% {
  print(i)
  foreach( e= 1:nrow(target), .combine = 'rbind',.packages = c("data.table", "raster")) %dopar% {
  print(e)
  result = cansee(Y_elev, origin[i,1:2], target[e,1:2], h1 = origin[i,3], h2 = target[e,3])
  result_row <- c(i,e, result)
  }
 # result_table <- rbind(result_table, result_row)
}

dim(ie_table) # 1,151,329 calculations take ~12 hours (10x less than for loop) 
saveRDS(ie_table, "../output_data/Yam_mnds_intervis.rds")
```
### Compare efficiency between for-loop and foreach
Let's run 4 * 4 (16) or/and 14*14 (~196) calculations using both a for loop and foreach to compare their relative speeds.

```{r compare-loop-foreach, eval = FALSE}
loop_table <- NULL
runtime1 <- system.time(
for (i in 1060:nrow(origin)){
  print(i)
  for (e in 1060:nrow(target)) {
    print(e)
    result = cansee(Y_elev, origin[i,1:2], target[e,1:2], h1 = origin[i,3], h2 = target[e,3])
    result_row <- cbind(i,e, result)
   loop_table <- rbind(loop_table, result_row)
    }
}
)

runtime2 <- system.time(
parallel_table <- foreach (i= 1060:nrow(origin), .combine = 'rbind',.packages = c("data.table", "raster")) %do% {
  print(i)
  foreach( e= 1060:nrow(target), .combine = 'rbind',.packages = c("data.table", "raster")) %dopar% {
  print(e)
  result = cansee(Y_elev, origin[i,1:2], target[e,1:2], h1 = origin[i,3], h2 = target[e,3])
  result_row <- c(i,e, result)
  }
}
)

dim(parallel_table)
dim(loop_table)
runtime1
runtime2
```

At 16 iterations (16 results), the foreach 'elapsed' time is half of the loop and 'user' time is 5x less.
> runtime1
   user  system elapsed 
   0.06    0.01    0.23 
> runtime2
   user  system elapsed 
   0.01    0.00    0.13 
   
At 196 iterations (196 results), the foreach (with 12 cores) 'elapsed' time is 1/3 of the for loop and 'user' time is 50x less    
> runtime1
   user  system elapsed 
   0.53    0.02    3.14 
> runtime2
   user  system elapsed 
   0.01    0.00    1.06
   
### Explore intervisibility results   
```{r results-test}
# Check the results of the initial intervisibility run
result_table <- readRDS("../output_data/result_table.rds")
result_table <- as.tibble(result_table)

head(result_table)
result_table <- result_table %>% 
  rename(i = V1, e = V2, visibility = V3)

# 10 most commanding mounds! i is the index, n is the number of visible features from it
result_table %>% 
  group_by(i) %>% 
  tally(visibility) %>% 
  arrange(desc(n)) # mounds no 333, 38 and 1 see the most other mounds (393-409)

# all the mounds visible from mound no. 333 (n = 409)
result_table %>% 
  filter(i == 333, visibility == 1) %>% 
  arrange(desc(e)) %>% 
  distinct(e) 
```

## Intervisibility of mounds in rasters with modelled vegetation
Rather than calculating LoS over bare-earth DEM, let's model it over surface where random stands of trees and scrub have been simulated. We load two rasters produced via 08_VisibilityProm.R:

1. DEM with vegetation modeled as random stands of trees of 10m height covering 50% of the extent of Yambol (randomly)
2. DEM with vegetation varying in height from 1-20m (normally distributed around a peak at 10) covering 50% of the extent of Yambol randomly

These DEMs are created by overlay of vegetation rasters and  Yambol elevation, with sum as the function on all values, increasing the terrain elevation by 10 - 20 m (producing Y_elev10 and Y_elev20grad respectively).

The hypothesis is that the vegetation will reduce the LoS among mounds that are at lower elevation differences or closer to one another. (Following Skov-Petersen)

```{r auto-foreach, eval = FALSE}
# install.packages("doParallel")
# install.packages("foreach")
library(foreach)
library(doParallel)
library(data.table)

detectCores() # 22 as some are multi-threaded
detectCores(logical = FALSE) # 16 real ones

cl <- makeCluster(12) # keep 4 cores so screensaver can run on W11
registerDoParallel(cl)

# Start experiment
origin <- cbind(st_coordinates(Yam_mnds), h = Yam_mnds$HeightMax)
target <- origin 

# nesting both i and e with foreach 
veg10_table <- foreach (i= 1:nrow(origin), .combine = 'rbind',.packages = c("data.table", "raster")) %do% {
  print(i)
  foreach( e= 1:nrow(target), .combine = 'rbind',.packages = c("data.table", "raster")) %dopar% {
  print(e)
  result = cansee(Y_elev10, origin[i,1:2], target[e,1:2], h1 = origin[i,3], h2 = target[e,3])
  result_row <- c(i,e, result)
  }
 # result_table <- rbind(result_table, result_row)
}
veg10_table$TRAPorigin <- Yam_mnds$TRAP[veg10_table$i]
veg10_table$TRAPtarget <- Yam_mnds$TRAP[veg10_table$e]

saveRDS(veg10_table, "output_data/veg10los.rds")

veg20_table <- foreach (i= 1:nrow(origin), .combine = 'rbind',.packages = c("data.table", "raster")) %do% {
  print(i)
  foreach( e= 1:nrow(target), .combine = 'rbind',.packages = c("data.table", "raster")) %dopar% {
  print(e)
  result = cansee(Y_elev20grad, origin[i,1:2], target[e,1:2], h1 = origin[i,3], h2 = target[e,3])
  result_row <- c(i,e, result)
  }
 # result_table <- rbind(result_table, result_row)
}
dim(veg20_table) # 1,151,329 calculations take ~12 hours (10x less than for loop) 

veg20_table$TRAPorigin <- Yam_mnds$TRAP[veg20_table$i]
veg20_table$TRAPtarget <- Yam_mnds$TRAP[veg20_table$e]

saveRDS(veg20_table, "output_data/veg20grad_los.rds")

```

### Intervisibility in a model with 0-20m vegetation
Let's explore what the random injection of 0-20m (10m mean height) vegetation into the elevation model. We anticipate that the overall intervisibility/ LoS will be reduced as a result of vegetation (as Skov-Petersen has demonstrated). 

```{r results-test-20}
# Check the results of the initial intervisibility run
result_table20 <- readRDS("output_data/veg20grad_los.rds")

head(result_table20)

# 10 most commanding mounds! i is the index, n is the number of visible features from it
result_table20 %>% 
  group_by(i) %>% 
  tally(visibility) %>% 
  arrange(desc(n)) # mounds no 333, 38 and 1 see the most other mounds (393-409)

# all the mounds visible from mound no. 38, TRAP 9044,(n = 291) Previously it was mound no.333  9412, which had 409 (now it has 279, 131 less than in bare earth model) 
result_table20 %>% 
  filter(i == 38, visibility == 1) %>% 
  arrange(desc(TRAPtarget)) %>% 
  distinct(TRAPtarget) 
```


### Most visible mound 9412 in veg20 gradual

```{r max-veg20}

# old most-seeing mound 9412 no. 333 now drops from 409 to 279)
result_table20 %>% 
  filter(i == 333, visibility == 1) %>% 
  arrange(desc(e)) %>% 
  distinct(TRAPtarget) %>% 
  pull() -> visiblefrom9412veg20  # list of TRAP numbers visible from 9412

# The new see-the-most mound is 9044 no. 38 (n = 291)
result_table20 %>% 
  filter(i == 38, visibility == 1) %>% 
  arrange(desc(e)) %>% 
  distinct(TRAPtarget) %>% 
  pull() -> visiblefrom9044veg20  # list of TRAP numbers visible from 9044


```

```{r most-commanding-200mnds}

los20veg_200 <- result_table20 %>% 
  group_by(TRAPorigin) %>% 
  tally(visibility) %>% 
  arrange(desc(n)) %>% # mounds no 333, 38 and 1 see the most other mounds but numbers 
  slice(1:200) %>% 
  pull(TRAPorigin)

```

```{r most-intervisible-veg20-map}

library(mapview)

result_table20 %>% #created from output_data 20los results_table 
  group_by(TRAPorigin) %>% 
  tally(visibility) %>% 
  arrange(desc(n)) %>% 
  slice(1:25) %>% 
  #rename(TRAP = TRAPorigin) %>% 
  #merge(Yam_mnds, by = c("TRAP")) #%>% 
  left_join(Yam_mnds, by = c("TRAPorigin"="TRAP")) %>%
  st_as_sf(crs = 32635) %>% 
  mapview(cex = "n" , zcol = "HeightMax") +mapview(Yam_mnds,cex = 0.1)
  mapview(Yam_mnds %>% dplyr::filter(TRAP%in%visiblefrom9044veg20), cex = 1)

```

In bare earth model, 9412 rules with line of sight to 409 other mounds. If we model random vegetation patches with mean 10m height, the mound with the greatest line of sight is 9044 a 6m tall mound near Botevo.


### Intervisibility in a model with 10m vegetation

Let's explore what happens after the random injection of exactly 10m high vegetation into the elevation model. We anticipate that the overall intervisibility/ LoS will be reduced as a result of vegetation (as Skov-Petersen has demonstrated), but perhaps not as much as in the 20m gradual model.

```{r results-test-veg10}
# Check the results of the initial intervisibility run
result_table10 <- readRDS("output_data/veg10los.rds")
result_table10 <- as_tibble(result_table10)

head(result_table10)
result_table10 <- result_table10 %>% 
  rename(i = V1, e = V2, visibility = V3)

# add TRAP IDs
result_table10$TRAPorigin <- Yam_mnds$TRAP[result_table10$i]
result_table10$TRAPtarget <- Yam_mnds$TRAP[result_table10$e]

```
### Most commanding mounds in veg10
```{r 200-most-see}
# 10 most commanding mounds! i is the index, n is the number of visible features from it
result_table10 %>% 
  group_by(i) %>% 
  tally(visibility) %>% 
  arrange(desc(n)) # mounds no 333, 38 and 1 see the most other mounds but numbers are even lower than in the 0-20 gradual height model (247-261), 333 leads again as it did in BEmodel.

# 200 most commanding mounds
los10veg_200 <- result_table10 %>% 
  group_by(TRAPorigin) %>% 
  tally(visibility) %>% 
  arrange(desc(n)) %>% # mounds no 333, 38 and 1 see the most other mounds but numbers 
  slice(1:200) %>% 
  pull(TRAPorigin)


# all the mounds visible from mound no. 333, TRAP 9412,(n = 261), lower than in veg20 (279) and less than in BEM(409).  
result_table10 %>% 
  filter(i == 333, visibility == 1) %>% 
  arrange(desc(e)) %>% 
  distinct(e) 

```


```{r most-intervisible-veg10-map}
result_table10 %>% 
  filter(i == 333, visibility == 1) %>% 
  arrange(desc(e)) %>% 
  distinct(TRAPtarget) %>% 
  pull() -> visiblefrom9412veg10

library(mapview)

result_table10 %>% 
  group_by(TRAPorigin) %>% 
  tally(visibility) %>% 
  arrange(desc(n)) %>% 
  slice(1:25) %>% 
  #rename(TRAP = TRAPorigin) %>% 
  #merge(Yam_mnds, by = c("TRAP")) #%>% 
  left_join(Yam_mnds, by = c("TRAPorigin"="TRAP")) %>%
  st_as_sf(crs = 32635) %>% 
  mapview(cex = "n" , zcol = "HeightMax") +mapview(Yam_mnds,cex = 0.1) #+
  #mapview(Yam_mnds %>% dplyr::filter(TRAP%in%visiblefrom9412veg10), cex = 1)
 # mapview(Yam_mnds %>% dplyr::filter(TRAP%in%los10veg_200), cex = 4) # 200 mounds with biggest fieldview in 10m vegetation model
```

In bare earth model, 9412 rules with line of sight to 409 other mounds. If we model random vegetation patches with mean 10m height, the mound with the greatest line of sight becomes 9044, a 6m tall mound near Botevo, with 291 visible mounds (with no limits on distance).

Let's now see what happens when we institute a series of buffers to model visual decline with distance/bad weather/low light.

Higuchi (Fig.2.12) lists that at noon, the visibility is:
45km in very clear weather, 40km in clear, 30km under cloud, 10km in drizzle and 6-4km in steady rain or snow and 0.6km under fog.
At dawn or dusk under the same condition, visibility drops to:
20km, 15km, 10km, 5-2km, and 0.3km.

## Intervisibility within 2, 5, 10km distance buffers
Let's see how distance buffers following different weather affect the amount of intervisible mounds.

Select from intervisible mounds those that intersect buffer of 1, 5,10, 20 and 45 km

```{r}
id <- 9418
buffer <- 5000
results <- mnd_vis 

visible_at_dist <- function(id, buffer){
  #create mound point
  moundbuff <- Yam_mnds %>%
    filter(TRAP == id) %>% 
    st_buffer(buffer)
  # create a vector of seen mounds id under ideal conditions
  seen_mnds <- results %>%
    filter(TRAPorigin == id, visibility == 1) %>%  
    select(TRAPtarget) %>% 
    pull()
  # filter which one of the seen mounds is visible within the distance 
  visible <- Yam_mnds %>% 
    filter(TRAP%in%seen_mnds) %>% 
    st_filter(moundbuff, .predicate = st_intersects) 
  # write interim result to a table
  vis_table <- cbind(id, buffer, targetid = visible$TRAP)
  # vis_table <- visible %>% 
  #   mutate(TRAPorigin = id, Buffer_km = buffer) %>% 
  #   relocate(c(TRAPorigin,Buffer_km), .before = TRAP) %>% 
  #   select(TRAPorigin, Buffer_km, TRAP) %>% st_drop_geometry()
  # attach interim table to mastertable
  #mapview(moundbuff, cex = 4) + mapview(visible)
  #final_table <- rbind(final_table, vis_table) 
}


# test
test <- visible_at_dist(9416,5000)
```

## eventually loop over different buffers to see which of the visible mounds falls to what distance band
```{r}
# create empty data container
final_table <- data.frame(data.frame(id = numeric(), buffer = numeric(), targetid = numeric()))

# vector of buffer distances
buffers <- c(1000,2500,5000,10000,20000,40000)


# test with 200 most commanding mounds los20veg200
los10veg_200
los20veg_200
losNOveg_200

# loop over the 200 mounds
for (id in losNOveg_200){
  print(id)
  for (buffer in buffers){
    result <- visible_at_dist(id, buffer)
    final_table <- rbind(final_table, result)
  } 
}  

#now I can calculate which mound sees most at shorter intervals
buffer_vis_table <- data.frame()
for (i in buffers) {
  bufvis <- final_table %>% 
  filter(buffer == i) %>% 
  group_by(id, buffer) %>% 
  tally() %>% 
  arrange(-n) %>% 
  ungroup() %>%  
  slice(1:20)  # the 20 most seeing mounds (with the highest number of seen objects)
  buffer_vis_table <- rbind(buffer_vis_table,bufvis)
}

# Summarize total number of mounds seen at each  buffer
summary <- buffer_vis_table %>%
  group_by(buffer) %>% 
  summarize(sum_seen_by200 = sum(n))

# Recalculate to mounds seen per sq km of radius (problematic because aggregate, but a clue about vegetation impact) WATCH WHICH ONE YOU RUN
mound_sums_per_buffer1 <- summary %>% 
  mutate(area_km2 = 3.14569*(buffer/1000)^2,
            mnd_km2 = sum/area_km2,
         type = "model10mveg")

mound_sums_per_buffer2 <- summary %>% 
  mutate(area_km2 = 3.14569*(buffer/1000)^2,
            mnd_km2 = sum/area_km2,
         type = "model20mveg")

mound_sums_per_buffer3 <- summary %>% 
  mutate(area_km2 = 3.14569*(buffer/1000)^2,
            mnd_km2 = sum/area_km2,
         type = "modelNOveg")
mound_sums_per_buffer3
mound_sums_per_buffer2
mound_sums_per_buffer
write_csv(rbind(mound_sums_per_buffer,
                mound_sums_per_buffer2,
                mound_sums_per_buffer3),
          "output_data/stuffseenby200veg.csv")
# nice, apply this to geneeral resuts and veg20losgrad to see how visibility declines  
```
If we take the 10 most commanding mounds across all buffers, we see that although area covered by radius grows geometrically `(pi*r^2)`, the number of visible features only grows linearly (*2)

CONTINUE # find the 10-20 most commanding mounds for each buffer zone and plot these


# and different visibility datasets
```{r}
library(tmap)
Yam_mnds %>% 
  full_join(buffer_vis_table , by = c("TRAP" ="id")) %>%  
  dplyr::filter(buffer == 40000) %>% 
  mapview(cex = "n")
  
```

### let's try with tmaps
```{r}
library(tmap)
tmap_mode(mode = "view")
Yam_mnds %>% 
  full_join(buffer_vis_table , by = c("TRAP" ="id")) %>%  
  dplyr::filter(!is.na(buffer)) %>% 
  tm_shape(.)+
  tm_facets(by = "buffer",
            ncol = 3)+
  tm_dots(size = "n")


```

## Which mounds appear constantly despite distance erosion and other limitations?
```{r}
bom_10 <- losNOveg_200[losNOveg_200 %in% los10veg_200]  # 108 are shared among the BOM and 10veg
veg10_20 <- los10veg_200[los10veg_200 %in% los20veg_200] # 91 are shared among 10 and 20veg
bom_10[bom_10%in%veg10_20] # 68 are shared among all lists  (persistently seeing mounds)

```



### Intervisibility with settlements

- test for visibility between mounds and major settlements - eg. Kabyle, Dadopara and further south Turkish sites.

Let's load archaeological settlements (BA-RM) known from Yambol
```{r load-yam-sites}
sites <- data.frame(name = c("Drazhevo","Kabyle", "Dadopara", "Stroino"),
                    lat = c(42.544722002584564,42.548058964016256,  42.221831,  42.292159),
                    long = c(26.44793978843489,26.48336749894564, 26.339603, 26.700300))

sites <- sites %>% 
  st_as_sf(coords = c("long", "lat"), crs = 4326)

y_sites <- read_sf("~/../Desktop/TRAP_Oxbow/YAM/YAM_scatterpoints.shp")
y_sites


library(mapview)
mapview(sites)+mapview(y_sites)
sites[2,]
```
```{r Kabyle-vis}
Kabyle <- sites %>% 
  filter(name == "Kabyle") %>% 
  st_transform(crs = 32635) %>% 
  st_coordinates()

result <- cansee(Y_elev, Kabyle , target[, 1:2], h1 = 2, h2 = target[,3])
  result_row <- c(i,e, result)

result_table <- NULL
for (e in 1:nrow(target)){print(e)
      result = cansee(Y_elev, Kabyle, target[e,1:2], h1 = 2, h2 = target[e,3])
    result_row <- cbind(e, result) 
    result_table <- rbind(result_table, result_row)
  }

head(result_table)
saveRDS(result_table, "output_data/Kabylesees.rds")

result_table <- readRDS("output_data/Kabylesees.rds")
result_table <- as_tibble(result_table)
result_table$TRAPtarget <- Yam_mnds$TRAP[result_table$e] 

Kabylesees <- as_tibble(result_table)

Kabylesees <- Kabylesees %>%
  filter(result == 1) 
 
Yam_mnds %>% 
  filter(TRAP %in% Kabylesees$TRAPtarget) %>% 
  mapview()
  
head(result_table)
```


## Visualizing intervisibility results


```{r most-intervisible-map}
# load data on visibility and major sites
mnd_vis <- readRDS("output_data/Yam_mnds_intervis.rds")

mnd_vis %>% 
  filter(i == 333, visibility == 1) %>% 
  arrange(desc(e)) %>% 
  distinct(TRAPtarget) %>% 
  pull() -> visiblefrom9412

# group by viewpoint, join to spatial points and plot on a map
library(mapview)

mnd_vis %>% 
  group_by(TRAPorigin) %>% 
  tally(visibility) %>% 
  arrange(desc(n)) %>% 
  slice(1:50) %>% 
  mutate(n_log = log10(n)) %>% 
  # join the summarized dataframe of 25 most visible mounds to mound spatial points
  left_join(Yam_mnds, by = c("TRAPorigin"="TRAP")) %>%
  # activate the geometry column with proper CRS
  st_as_sf(crs = 32635) %>% 
  # plot the 25 most visible mounds, varying their radius by number of visible features from this location and varying color by mound height
  mapview(cex = "n" , zcol = "n_log") + 
  # show other mounds in the background as small dots
  mapview(Yam_mnds,cex = 0.1) 
  # differentiate mounds visible from 9412 (the most visible single features)
  #mapview(Yam_mnds %>% dplyr::filter(TRAP%in%visiblefrom9412), cex = 1)

mapview(Yam_mnds[Yam_mnds$TRAP %in% visiblefrom9412,])
```
### most visible
### Visibility lines for 9412
```{r most-commanding200-inBOM}
losNOveg_200 <- mnd_vis %>% 
  group_by(TRAPorigin) %>% 
  tally(visibility) %>% 
  arrange(desc(n)) %>% # mounds no 333, 38 and 1 see the most other mounds but numbers 
  slice(1:200) %>% 
  pull(TRAPorigin)


```


```{r linestrings-9412}
# Test linestring creation from 9412 and its 409 visible mounds. The coords object needs to be a matrix, but to cbind one to many coordinates the component columns need to be dataframes.

mostvisiblemnd <- visiblefrom9044veg20 # fill in model (veg10, veg20, none) for which we are calculating the lines of sight to the most visible mound

coords <- as.matrix(cbind(as.data.frame(st_coordinates(Yam_mnds %>% filter(TRAP == 9412))),
                as.data.frame(st_coordinates(Yam_mnds %>% filter(TRAP %in% mostvisiblemnd)))))

lines_sm <-  st_sfc(
     lapply(1:nrow(coords),
           function(i){
             st_linestring(matrix(coords[i,],ncol=2,byrow=TRUE))
           }))

st_crs(lines_sm) <- st_crs(Yam_mnds)


library(mapview)
mapview(lines_sm)+ 
  mapview(Yam_mnds %>% filter(TRAP %in% mostvisiblemnd))
plot(lines_sm); plot(Yam_mnds %>% filter(TRAP %in% mostvisiblemnd) %>% st_geometry, add =T)
# https://stackoverflow.com/questions/65498300/how-to-efficiently-create-linestrings-from-points
# https://stackoverflow.com/questions/58150279/plotting-lines-between-two-sf-point-features-in-r

```

The linestrings for 9412 seem to reach the national border, which is hard to believe... 
What is going on here? Let's check the prominence of the points first and raster profiles second.


## Raster profile
```{r}
profile <- function(IDorigin, IDtarget){
     # must be an elevation raster
  # library(raster)
  # if(!exists(Y_elev)){
  #    Y_elev <- raster("../output_data/large/Yelev32635.tif")}
     # must be a simple feature with same CRS as raster
  library(sf) 
  # target just as a dataframe of x, y, WITHOUT height
  origin <- Yam_mnds %>%  
         filter(TRAP == IDorigin) %>% 
          st_coordinates()  
  target <- Yam_mnds %>% 
         filter(TRAP == IDtarget) %>%
          st_coordinates()
  # target needs to be a dataframe of x, y, and height
  # origin <- data.frame(Yam_mnds %>%  
  #        filter(TRAP == origin) %>% 
  #         st_coordinates(), h= Yam_mnds$HeightMax[Yam_mnds$TRAP == origin])
  #   target <- data.frame(Yam_mnds %>% 
  #        filter(TRAP == target) %>% 
  #         st_coordinates(), h= Yam_mnds$HeightMax[Yam_mnds$TRAP == target])
  testprofile <- rasterprofile(Y_elev,  origin, target)
  x = c(0,nrow(testprofile))
  y = testprofile$z[c(1,nrow(testprofile))]

    # Plot profile with start and end point elevation
  plot(testprofile$z, type = "l", 
       main = paste0("Terrain profile between mounds ", IDorigin, " and ", IDtarget),  
       ylab = "Elevation(m)", xlab = "Mound distance in pixels (*30m)"); 
  lines(x,y, col = "red", lwd = 2)
  
}

 # these two points are at opposite sides of the region
profile(9412, 8242)
profile(9044, 8251 )
# Plot
plot(testprofile$z, type = "b"); 

```
## 8007
```{r}
mnd_vis %>% 
  filter(TRAPorigin == 8007, visibility == 1) %>% 
  arrange(desc(e)) %>% 
  distinct(TRAPtarget) %>% 
  pull() -> visiblefrom8007NOveg  # list of TRAP numbers visible from 9044

result_table10 %>% 
  filter(TRAPorigin == 8007, visibility == 1) %>% 
  arrange(desc(e)) %>% 
  distinct(TRAPtarget) %>% 
  pull() -> visiblefrom8007veg10  # list of TRAP numbers visible from 9044

result_table20 %>% 
  filter(TRAPorigin == 8007, visibility == 1) %>% 
  arrange(desc(e)) %>% 
  distinct(TRAPtarget) %>% 
  pull() -> visiblefrom8007veg20  # list of TRAP numbers visible from 9044

mostvisiblemnd <- visiblefrom8007veg20 # fill in model (veg10, veg20, none) for which we are calculating the lines of sight to the most visible mound

coords <- as.matrix(cbind(as.data.frame(st_coordinates(Yam_mnds %>% filter(TRAP == 8007))),
                as.data.frame(st_coordinates(Yam_mnds %>% filter(TRAP %in% mostvisiblemnd)))))

lines_sm <-  st_sfc(
     lapply(1:nrow(coords),
           function(i){
             st_linestring(matrix(coords[i,],ncol=2,byrow=TRUE))
           }))

st_crs(lines_sm) <- st_crs(Yam_mnds)


library(mapview)
mapview(lines_sm)+ 
  mapview(Yam_mnds %>% filter(TRAP %in% mostvisiblemnd))
```

## Kabyle
```{r}
Kabyle <- Kabyle %>% 
  as_tibble() %>% 
  st_as_sf(coords = c("X", "Y"), crs = 32635)  
Kabylesees

mostvisiblemnd <- Kabylesees$TRAPtarget # fill in model (veg10, veg20, none) for which we are calculating the lines of sight to the most visible mound

mostvisiblemnd <- visiblefrom8007veg20 # fill in model (veg10, veg20, none) for which we are calculating the lines of sight to the most visible mound

coords <- as.matrix(cbind(as.data.frame(st_coordinates(Kabyle)),
                as.data.frame(st_coordinates(Yam_mnds %>% filter(TRAP %in% mostvisiblemnd)))))

lines_sm <-  st_sfc(
     lapply(1:nrow(coords),
           function(i){
             st_linestring(matrix(coords[i,],ncol=2,byrow=TRUE))
           }))

st_crs(lines_sm) <- st_crs(Yam_mnds)


library(mapview)
mapview(lines_sm)+ 
  mapview(Yam_mnds %>% filter(TRAP %in% mostvisiblemnd))

```

## Chronology
```{r mnds-google}
library(googlesheets4)
gs4_deauth()
mnd_izv <- read_sheet("https://docs.google.com/spreadsheets/d/1wOxbKVHGNHox4h86Z5ZXXADUubMJofC1KqDLMLT13a8/edit#gid=369527838", 
                      range = "General",
                      col_types = 'nccccnnccncccccccccccc')

burial_izv <- read_sheet("https://docs.google.com/spreadsheets/d/1wOxbKVHGNHox4h86Z5ZXXADUubMJofC1KqDLMLT13a8/edit#gid=369527838", 
                      range = "BurialAttributes",
                      col_types = 'ncnccccnccccccccccccccccccccnncccc')

mnd_aor <- read_sheet("https://docs.google.com/spreadsheets/d/1cx0nntcCLgrwQvCvvIYkjFJ-TnoJ0QRJfTgrf2qEEv4/edit#gid=1795672216", 
                      range = "GeneralSpatial",
                      col_types = 'nccnnncccc')

burial_aor <- read_sheet("https://docs.google.com/spreadsheets/d/1cx0nntcCLgrwQvCvvIYkjFJ-TnoJ0QRJfTgrf2qEEv4/edit#gid=1795672216", 
                      range = "BurialAttributes",
                      col_types = 'nccccnccccccccccccccnnc')

glimpse(mnd_izv)
glimpse(burial_izv)
glimpse(mnd_aor)
glimpse(burial_izv)

Yam_aor <- mnd_aor %>% 
  full_join(burial_aor, by = "MoundID") %>% 
  filter(!is.na(Long)) %>% 
  st_as_sf(coords = c("Long", "Lat"), crs = 4326) %>% 
  filter(Region == "Yambol") 
 
unique(Yam_aor$MoundID) # 40 mounds

Yam_izv <- mnd_izv %>% 
  full_join(burial_izv, by = "MoundID") %>% 
  filter(!is.na(Long)) %>% 
  st_as_sf(coords = c("Long", "Lat"), crs = 4326) %>% 
  filter(Region == "Yambol")
unique(Yam_izv$MoundID) # 10 mounds
Yam_izv$StartDate

mapview(Yam_aor, zcol = "StartDate") + mapview(Yam_izv, zcol = "StartDate") +
mapview(Yam_mnds %>% filter(prom250mbuff >75))
```




## Prominence point map
We have mound-centereed indeces of ruggedness and topographic prominence, which should align with the objective/dry measure of visibility above.
The measures are at different scales. Prominence in 250m buffer specifies how big a fraction (in percentages) of the surrounding area is visible to a person on the mound.
TPI is measured from on a scale from -3 to 2 , with most values between 0 and 1. TRI is between 0 and 4, with median at 1. Rough measure goes from 0 to 15, with mean at 3.7.
```{r}
# various measures of prominence
summary(Yam_mnds[, c("prom250mbuff", "TPI", "TRI", "rough")])


# prominence in 250m buffer
hist(Yam_mnds$prom250mbuff)
Yam_mnds %>% 
  filter(prom250mbuff > 85) %>% 
  mapview(zcol = "prom250mbuff") # 9412 not present!

# TPI
hist(Yam_mnds$TPI)
Yam_mnds %>% 
  filter(TPI >= 1) %>% 
  mapview(zcol = "TPI")# 9412 not present!

# TRI
hist(Yam_mnds$TRI)
Yam_mnds %>% 
  filter(TRI >= 2.53) %>% 
  mapview(zcol = "TRI")# 9412 not present!

# rough
hist(Yam_mnds$rough)
Yam_mnds %>% 
  filter(rough >= 7) %>% 
  mapview(zcol = "rough")# 9412 not present!
```


### BA mound subgroup
to keep things simple
```{r BA-faims, include = FALSE}
# Filter out the BA mounds from the FAIMS data
bronze <- c(6009,8007,6010,8345,8346,8357, 8502, 8351, 9257)
BA <- Yam_mnds %>% 
  filter(TRAP %in% bronze)

# Add literature dimensions to mound 8345, which was excavated before survey 
BA$DiameterMax[5] <- 37.5
BA$HeightMax[5] <- 3.2

```


```{r BA-cansee}
mapview(BA[c(3,6,5),])
m8007 <- c(st_coordinates(BA)[3,],h= BA$HeightMax[3])
m8346 <- c(st_coordinates(BA)[6,],h= BA$HeightMax[6])
m8502<- c(st_coordinates(BA)[7,],h= BA$HeightMax[7])
m8345 <- c(st_coordinates(BA)[5,],h= BA$HeightMax[5])
Y_elev

test_see <- cansee(Y_elev, m8345, m8346, h1 = m8345[3]  , h2 = m8346[3])
test_see

```
## Linestring creation

Let's first see if we an create linestrings to later signal which mounds can *see* each other. To do that we need two columns of coordinates: the origin and the target. Here we explore the intervisibility of the seven BA mounds
```{r linestring}
# Test linestring creation from one origin and multiple target points. The coords object needs to be a matrix, but to cbind one to many coordinates the component columns need to be dataframes.

coords <- as.matrix(cbind(as.data.frame(st_coordinates(BA[2])),
                as.data.frame(st_coordinates(BA))))

lines_sm <-  st_sfc(
     lapply(1:nrow(coords),
           function(i){
             st_linestring(matrix(coords[i,],ncol=2,byrow=TRUE))
           }))

st_crs(lines_sm) <- st_crs(BA)


library(mapview)
mapview(lines_sm)+mapview(BA)

# https://stackoverflow.com/questions/65498300/how-to-efficiently-create-linestrings-from-points
# https://stackoverflow.com/questions/58150279/plotting-lines-between-two-sf-point-features-in-r
```

### Visualizing the terrain profile
```{r raster-profile}
# Create a terrain profile and plot it to check potential intervisibility
testprofile <- rasterprofile(Y_elev,  m8007, m8346)
testprofile <- rasterprofile(Y_elev,  m8007, m8345)
testprofile <- rasterprofile(Y_elev,  st_coordinates(Yam_mnds[1,]), st_coordinates(Yam_mnds[333,]))

# Plot
plot(testprofile$z, type = "b"); 

# create the connecting line between origin and destination points
x = c(0,nrow(testprofile))
y = testprofile$z[c(1,nrow(testprofile))]

# Plot profile with start and end point elevation
plot(testprofile$z, type = "l", main = "Terrain profile between mounds 8345 and 8346", 
     ylab = "Elevation(m)", xlab = "Mound distance (*30m)"); 
lines(x,y, col = "red", lwd = 2)

```

Some of the mounds (Drazhevo and 8346) can not see each other, because the one on top of the Zaichi Vrah is obscured by the hill and not aimed at the mounds to the SE.  Others with clear line of sight can see one another.
Let's test the intervisibility between these three and other mounds in the surroundings from the Yambol Mounds dataset. We will use the viewTo() function and the starting coordinates of the three mounds.

```{r view-to-8007, warning=FALSE}
# First we need to convert the Yambol mounds to a matrix of coordinates
coords <- data.frame(st_coordinates(Yam_mnds),h=Yam_mnds$HeightMax)
missingHeight <- which(is.na(coords$h))
mnd_missH <- Yam_mnds$TRAP[missingHeight]

coords <- coords %>% 
  filter(!is.na(h))
coords_m <- as.matrix(coords)

# viewTo() working with 3 points
testview <- viewTo(Y_elev, m8007, rbind(m8346,m8345))

# with all points (uncomment if you wish to regenerate!)
# testview <- viewTo(Y_elev, m8007, coords_m,m8007[3], coords_m[,3])
# length(testview)


# # See which ones can see 8007
# '%nin%' = Negate('%in%')
# viewTo_8007 <- Yam_mnds %>% 
#   filter(TRAP %nin% mnd_missH) %>%   # must first filter away missing heights as I did with coordinates
#   filter(testview) # use the logical vector to filter the Yam_mnds rows

# 289 mounds can see mound 8007
# saveRDS(viewTo_8007, "../output_data/MoundsVisfrom8007.rds")
viewTo_8007 <- readRDS("../output_data/MoundsVisfrom8007.rds")
```

View these 289 mounds visible from 8007
```{r plot-visible}
# Test linestring creation from one origin and multiple target points. The coords object needs to be a matrix, but to cbind one to many coordinates the component columns need to be dataframes.
coords <- as.matrix(cbind(rep(m8007[1], 289),rep(m8007[2], 289), st_coordinates(viewTo_8007)))

lines <-  st_sfc(
     lapply(1:nrow(coords),
           function(i){
             st_linestring(matrix(coords[i,],ncol=2,byrow=TRUE))
           }))

st_crs(lines) <- st_crs(BA)

library(mapview)
library(leafsync)
m1 <- mapview(lines) + mapview(viewTo_8007, cex = 4) + mapview(BA[3,], cex = 10)
m2 <- mapview(viewTo_8007, zcol = "prom250mbuff")
sync(m1,m2)
```




## Intervisibility of Bronze Age mounds in Yambol (not sure if needed)
In this proof of concept we explore the visibility of BA mounds from other BA mounds, for simplicity's sake. 

BA mounds comprise surveyed and excavated mounds ('bronze') and excavated prior to survey ('extrasf') supplied by Toshko.

Excavated prior:
Drazhevo - Height: 2.7 m; Diameter: 37.0 m; Land: pasture on the rocky area
Straldzha - Height: 4.0 m; Diameter: 50.0 m; Land: wines and pasture
Popovo - Height: 6.8 m; Diameter: 48.0 m; Land: wines and pasture

8345 Mogila village - Height: 3.2 m; Diameter: 37.6 m; Land: pasture on the rocky area 


```{r BA-faims, include = FALSE}
# Filter out the BA mounds from the FAIMS data
bronze <- c(6009,8007,6010,8345,8346,8357, 8502, 8351, 9257)
BA <- Yam_mnds %>% 
  filter(TRAP %in% bronze)

# Add literature dimensions to mound 8345, which was excavated before survey 
BA$DiameterMax[5] <- 37.5
BA$HeightMax[5] <- 3.2

```

```{r BA-Todor, include = FALSE}
# Add the few mounds excavated before TRAP coverage >> these should be pulled out of the AOR dataset to get the Height, Diameter and other dimensions..
extras <- data.frame(place = c("Drazhevo", 
                           "Straldzha",
                           "Popovo"),
                 Easting=c(455442,479522,479391),
                 Northing=c(4710935,4715730,4668279),
                 HeightMax=c(2.7,4.0,6.8),
                 DiameterMax = c(37,50,48),
                 LU_Around = c("Pasture","Perennial","Perennial"),
                 Type = c("Burial Mound","Burial Mound","Burial Mound"),
                 Source = c("Excavation","Excavation","Excavation"))
extrasf <- st_as_sf(extras, coords =c("Easting", "Northing"), crs = 32635)

```

```{r BA-combo, include = FALSE}
# Combine both BA datasets and spatialize
library(plyr)
BApoints <- rbind.fill(BA, extrasf)
BAcoords <- rbind(st_coordinates(BA) ,st_coordinates(extrasf))
BApoints <- cbind(BApoints, BAcoords)
BApoints <- st_as_sf(BApoints, coords = c("X", "Y"), crs = 32635)
```
