---
title: "Intervisibility Assessment"
output: html_document
date: "2024-06-20"
---
## Purpose and prerequisites
This script assesses  intervisibility (line of sight) results, starting here with burial mounds in the Yambol Region within ASTER elevation raster (30m spatial resolution) looking inside and outside the region. 

This markdown guides you to: 

1. Visualize intervisibility for calculated features (here: mounds in Yambol). To visualize, you can skip directly to section "Visualizing Intervisibility".
3. TBD: intervisibility between mounds and settlements , started but needs completion

The script basically works, but depends on the datasetes

The proof of concept is done on BA mounds and to replicate it you might need: 
- to run 09 BA mounds to have the necessary libraries and digital objects

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE)
library(sf)
library(raster)
library(tidyverse)
library(mapview)

```

## Load Data

```{r}
#YT_elev <- raster("data/large/YT_elev32635.tif") # vertical
Y35 <- raster("data/large/Y_elev_ext.tif") # horizontal

# region admin
region <- read_sf("data/YamRegion.shp")

# Yambol mounds
Yam_mnds <- readRDS("data/Yam_dd_mnds.rds") # 1073 features
mm_Yext <- read_sf("output_data/Y25_mapmounds.shp")

# Intervisibility
Y_in <- readRDS("output_data/Yam_mnds_intervis.rds")
Y_out <- readRDS("output_data/Yam_ext_mnds_intervis.rds")
```

## Final visibility review
```{r in_ext-vis}
Y_vis <- Y_in %>% rbind(Y_out)
all_vis_sorted <- Y_vis %>% 
  group_by(TRAPorigin) %>% 
  tally(visibility) %>% 
  arrange(desc(n))

# the winners of outside-looking mounds
Y_out %>% 
  group_by(TRAPorigin) %>% 
  tally(visibility) %>% 
  arrange(desc(n)) %>% slice(1:10) %>% pull(TRAPorigin) -> outside10
Y_in %>% 
  group_by(TRAPorigin) %>% 
  tally(visibility) %>% 
  arrange(desc(n)) %>% slice(1:10) %>% pull(TRAPorigin) -> inside10
outside10[which(outside10%in%inside10)]

Y_vis %>% 
  group_by(TRAPorigin) %>% 
  tally(visibility) %>% 
  arrange(desc(n))%>% slice(1:10) %>% pull(TRAPorigin) -> all10

outside10[which(outside10%in%inside10)]
all10[all10%in%outside10]


```


## Tables

```{r}
Y_out %>% 
  group_by(TRAPorigin) %>% 
  tally(visibility) %>% 
  arrange(desc(n)) %>% pull(n) %>% hist(main = "Number of intervisible mounds")

```
## Histograms

### Simple of mounds seen, totals and log
```{r}

Y_vis %>%
group_by(TRAPorigin) %>%
tally(visibility) %>%
arrange(desc(n)) %>%
ggplot()+
geom_histogram( aes(x = n)) +
theme_bw()+
labs( x = "Number of seen mounds ")

Y_vis %>%
group_by(TRAPorigin) %>%
tally(visibility) %>%
arrange(desc(n)) %>%
mutate(n_log = log10(n)) %>%
ggplot()+
geom_histogram( aes(x = n_log)) +
theme_bw()+
labs( x = "Number of seen mounds (log10)")


```

### Classified in intervisibility bins

While I can cacluate how many mounds can each mound see (in theory), it's also good to assess mound belonging into one of three logarithmic classes of intervisiblity, seeing 1 to 10 mounds, 10 to 100 mounds and 100 to 1000 mounds. And eventually assess how membership in these classes changes if we use vegetation models.
```{r}
# histogram
Y_vis %>%
  group_by(TRAPorigin) %>%
  tally(visibility) %>%
  arrange(desc(n)) %>%
  mutate(n_log = log10(n)) %>%
  mutate(n_class = case_when(n_log < 1 ~ "1 to 10",
                              n_log < 2 ~ "10 to 100",
                              n_log < 3 ~ "100 to 1000")) %>%
  ggplot() +
  geom_histogram(aes(x = n_class), stat="count") +
  theme_bw()+
  labs(x = "How many mounds can be seen from this location?")  # add text to the columns?

# count the intervisibility class contents 
Y_vis %>%
  group_by(TRAPorigin) %>%
  tally(visibility) %>%
  arrange(desc(n)) %>%
  mutate(n_log = log10(n)) %>%
  mutate(n_class = case_when(n_log < 1 ~ "1 to 10",
                            n_log < 2 ~ "10 to 100",
                            n_log < 3 ~ "100 to 1000")) %>%
  group_by(n_class) %>%
  tally()

# count individual mound intervisibility
Y_vis %>%
  group_by(TRAPorigin) %>%
  tally(visibility) %>%
  arrange(desc(n))
```


## Maps
```{r}
all_vis_sorted %>% 
  mutate(log_vis = log10(n)) %>% 
  slice(1:100) %>% 
  left_join(Yam_mnds, by = c("TRAPorigin"="TRAP")) %>%
  st_as_sf(crs = 32635) %>% 
  mapview(cex = "log_vis" , zcol = "n") + mapview(Yam_mnds,cex = 0.1) + mapview(mm_Yext, cex = 0.1) + mapview(region, alpha = 1)
```


## Vegatation models and Intervisibility 

### model with 10m vegetation
Let's explore what the random injection of 10m (10m mean height) vegetation into the elevation model. We anticipate that the overall intervisibility/ LoS will be reduced as a result of vegetation (as Skov-Petersen has demonstrated). 

```{r results-test-10}
# Check the results of the initial intervisibility run
result_table10 <- readRDS("output_data/veg10los.rds")
vis10veg_sorted <-result_table10 %>% 
  group_by(TRAPorigin) %>% 
  tally(visibility) %>% 
  arrange(desc(n))
vis10veg_sorted

```


### model with 0-20m vegetation
Let's explore what the random injection of 0-20m (10m mean height) vegetation into the elevation model. We anticipate that the overall intervisibility/ LoS will be reduced as a result of vegetation (as Skov-Petersen has demonstrated). 

```{r results-test-20}
# Check the results of the initial intervisibility run
result_table20 <- readRDS("output_data/veg20grad_los.rds")
vis20veg_sorted <-result_table20 %>% 
  group_by(TRAPorigin) %>% 
  tally(visibility) %>% 
  arrange(desc(n))
vis20veg_sorted
```

### Assessing decline in field of view with vegetation
```{r}
intervis <- all_vis_sorted %>% 
  full_join(vis10veg_sorted, by = "TRAPorigin") %>% 
  full_join(vis20veg_sorted, by = "TRAPorigin") %>% 
  rename(vis_max = n.x, vis_10 = n.y, vis_20 = n)

top10 <- intervis %>% 
  arrange(desc(vis_max)) %>% slice(1:10) %>% pull(TRAPorigin)

top10in10 <- intervis %>% 
  arrange(desc(vis_10)) %>% slice(1:10) %>% pull(TRAPorigin)
top10[which(top10%in%top10in10)]
  
top10in20 <- intervis %>% 
  arrange(desc(vis_20)) %>% slice(1:10) %>% pull(TRAPorigin)
top10[which(top10%in%top10in10)] %in% top10[which(top10%in%top10in20)]

intervis %>% 
  mutate(decline10 = round(100-(vis_10/vis_max*100),2),
         decline20 = round(100-(vis_20/vis_max*100),2), 
         diff = decline10-decline20) %>% 
  #filter(decline10>90)  # 418 mounds lose visibility if in veg 10 model
  filter(decline20>90) # 206 mounds lose visibility if in veg 20 model
         # avgdec10 = mean(decline10),65%
         # avgdec20 = mean(decline20)) 60%
```

## Factoring vegetation models into intervisibility class membership

